|

### 2025-07-07 output/IMPS-2025-slideshow.qmd
Translate pending verbatim in slides

# Probability axes:
PROB_AXIS <- seq(-3.3, 3.3, by = .01)
PROB_AXIS_2D <- seq(-1.5, 1.5, by = 0.05) # Axis for the probability contour plots
```

@@

M2PL model (mckinley_extension_1983):

::::: columns
::: column
```{r mirt-model-irs-3d}
#| fig-width:  5
#| fig-height: 5
PROB_AXIS_2D <- seq(-3, 3, by = .1)

mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

@@

contour_plot
```

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                                        +
  geom_line(linewidth = LINE_WIDTH)                        +
  labs(x = "Responsabilidad", y = "P(Y~i~ = 1)")           +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                                        +
  scale_x_continuous(
    minor_breaks = NULL,
    expand = expansion()
  )                                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.x     = element_text(margin = margin(t = 1, unit = "lines")),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
    # The following is necessary to avoid showing the axis breaks, as dropping
    #   them prevents the axis line from being drawn (see
    #   https://github.com/tidyverse/ggplot2/issues/2983)
    axis.text.x.bottom = element_blank(),
    axis.ticks.length.x.bottom = unit(0, units = "cm"),
    panel.grid.major.x = element_blank()
  )
```
:::

@@

![Ítems de "ACT Mathematics Usage Test"](../www/Fig_1-6_Contemporary_Psychometrics.png){fig-align="left"}
:::
::::::

@@

::::: columns
::: column
![Un cubo rojo en perspectiva isométrica](../www/slideshow-assets/isometric-cube.jpg){fig-align="center"}
:::

@@

::: r-center
![Ítems de "ACT Mathematics Usage Test"](../www/Fig_1-6_Contemporary_Psychometrics.png){fig-align="center" height="400px"}
:::

@@

¿Cuál es el sistema de coordenadas de *θ* para que *θ\** sea "esférica"?

@@

::: fragment
[Respuestas en 3, 2, 1...](https://wall.sli.do/event/dPunB4BMx3S8hfSjaFSuFh?section=ae77dd00-34fb-4600-9677-337a65b04b83)
:::

@@

```{r item-plot-out}
#| fig-cap:   Espacio test ortogonal (izquierda) y oblicuo (derecha)
#| fig-align: center
#| fig-asp:     .5
#| fig-width: 13

@@

Parámetros multidimensionales generalizados

-   A ítems con monotonía constante

-   A Ejes oblicuos

@@

::: fragment
Distinción espacio latente - espacio test
:::

>>>>

# Probability axes:
PROB_AXIS <- seq(-3.3, 3.3, by = .01)
PROB_AXIS_2D <- seq(-3, 3, by = 0.05) # Axis for the probability contour plots
```

@@

M2PL model [@mckinley_extension_1983]: $P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)$

::::: columns
::: column
```{r mirt-model-irs-3d}
#| fig-width:  5
#| fig-height: 5
mirt_item <- tribble(
  ~item, ~a_1, ~a_2, ~d,
  1,     1,    0.5,  -0.5
)

@@

contour_plot
```
:::
:::::

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                                        +
  geom_line(linewidth = LINE_WIDTH)                        +
  labs(x = "Conscientiousness", y = "P(Y~i~ = 1)")           +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                                        +
  scale_x_continuous(
    minor_breaks = NULL,
    expand = expansion()
  )                                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.x     = element_text(margin = margin(t = 1, unit = "lines")),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
    # The following is necessary to avoid showing the axis breaks, as dropping
    #   them prevents the axis line from being drawn (see
    #   https://github.com/tidyverse/ggplot2/issues/2983)
    axis.text.x.bottom = element_blank(),
    axis.ticks.length.x.bottom = unit(0, units = "cm"),
    panel.grid.major.x = element_blank()
  )
```
:::

@@

!["ACT Mathematics Usage Test" items](../www/Fig_1-6_Contemporary_Psychometrics.png){fig-align="left"}
:::
::::::

@@

::::: columns
::: column
![A red cube in isometric perspective](../www/slideshow-assets/isometric-cube.jpg){fig-align="center"}
:::

@@

::: r-center
!["ACT Mathematics Usage Test" items](../www/Fig_1-6_Contemporary_Psychometrics.png){fig-align="center" height="400px"}
:::

@@

¿What is the coordinate system of *θ* to make *θ\** "spherical"?

@@

::: fragment
[Answers in 3, 2, 1...](https://wall.sli.do/event/dPunB4BMx3S8hfSjaFSuFh?section=ae77dd00-34fb-4600-9677-337a65b04b83)
:::

@@

```{r item-plot-out}
#| fig-cap:   Orthogonal (left) and oblique (right) test space
#| fig-align: center
#| fig-asp:     .5
#| fig-width: 13

@@

Generalization of multidimensional parameters

-   To constant-monotony items

-   To oblique axes

@@

::: fragment
Distinction latent space - test space
:::
